On the probability of finding a nonphysical solution through shadowing

### 2020-10-29 main.tex
Add AMM reference

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical solutions play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations for which the physical 
distribution is drastically different from the physical distribution 
of the original governing equation? Sauer \cite{sauer} has shown 
examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor. The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}). Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}.

Both the nonphysicality of shadowing solutions, and the existence of extremely non-smooth statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results. This is because shadowing-based methods compute the sensitivities of ensemble statistics along shadowing solutions, but these may not be physical solutions that reproduce ensemble statistics. The error in shadowing sensitivities has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. 
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{osc_tent_0.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_3.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_3.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_6.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
    \caption{
    The effect of increasing $n$ on the probability distribution associated to the plucked tent map. The left column shows the plucked tent map at 
    different values of $s$ and $n = 0$ (top), $n=3$ (middle), $n=6$ (bottom). The original tent map is at $s=0$ on each plot.
    The right column shows the stationary, physical probability distribution of the plucked tent map at $s=0.1$ and $n = 0$ (top), $n=3$ (middle), $n = 6$ (bottom).}
    \label{fig:pluckedTentMap}
\end{figure}

@@

In this paper, we have constructed several counterexamples that 
dispel the notion that shadowing solutions are physical solutions, i.e., that their 
statistical distribution is the same as that typically observed for almost every solution 
of a governing equation. The existence of long-time shadowing solutions \cite{grebogi} has historically been used to address the issue of whether numerical simulations, which are 
perturbed solutions, represent the true physics implied by the governing equation. In light of the evidence in this paper, we must reopen this issue. Even when numerical simulations are 
shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term or ensemble behavior of the physical system. The nonphysicality of shadowing solutions also indicates that shadowing-based methods \cite{qiqi-lss} can lead to incorrect values of sensitivities of statistics to parameter changes.

>>>>

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical solutions play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations for which the physical 
distribution is drastically different from the physical distribution 
of the original governing equation? Sauer \cite{sauer} has shown 
examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor; other works \cite{cns} further analyze the unreliability due to double-precision arithmetic of statistical measures such as power spectra and correlation functions. The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}). Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}.

Both the nonphysicality of shadowing solutions, and the existence of extremely non-smooth statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results. This is because shadowing-based methods compute the sensitivities of ensemble statistics along shadowing solutions, but these may not be physical solutions that reproduce ensemble statistics. The error in shadowing sensitivities has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. 
\begin{figure}
    \centering
    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_0.png}
     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_0.png}
    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_3.png}
     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_3.png}
    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_6.png}
     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_6.png}
    \caption{
    The effect of increasing $n$ on the physical probability distribution associated to the plucked tent map. The left column shows the plucked tent map at 
    different values of $s$ and $n = 0$ (top), $n=3$ (middle), $n=6$ (bottom). The original tent map is at $s=0$ on each plot.
    The right column shows the stationary, physical probability distribution of the plucked tent map at $s=0.1$ and $n = 0$ (top), $n=3$ (middle), $n = 6$ (bottom).}
    \label{fig:pluckedTentMap}
\end{figure}

@@

In this paper, we have constructed several counterexamples that 
dispel the notion that shadowing solutions are physical solutions, i.e., that their 
statistical distribution is the same as that typically observed for almost every solution 
of a governing equation. The existence of long-time shadowing solutions \cite{grebogi} has historically been used to address the issue of whether numerical simulations, which are 
perturbed solutions, represent the true physics implied by the governing equation. In light of the evidence in this paper, we must reopen this issue. Even when numerical simulations are 
shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term or ensemble behavior of the physical system. The nonphysicality of shadowing solutions also indicates that shadowing-based methods can lead to incorrect values of sensitivities of statistics to parameter changes.
